---
title: "Installing Packages"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{install_the_package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

## Installation from GitHub  

`charite` is a minimal R package that extends `gglot2` functionalities with a custom theme and standardized colour palettes styled to match the  visual identity of [Charité – Universitätsmedizin Berlin](https://marke.charite.de/d/Y3FxSwD6Tz3a). The package was developed by [Johannes Julius Mohn](https://cognition.maxplanckschools.org/en/doctoral-candidates/johannes-j-mohn).

You can install the package directly from GitHub. 

```{r install from GitHub}
# install.packages("devtools") # install `devtools` if not yet available
devtools::install_github("johannesjuliusm/charite")
```

## Installation from source  

If you are working on a computer with proxies, installation requires a little workaround. A proxy is an intermediary server that handles your internet traffic. This setup is common in corporate, university, or government networks where internet access is controlled or filtered.  

Go to the [GitHub page](https://github.com/johannesjuliusm/charite) and download the package tarball (.tar.gz) from the "Releases" tab on the right hand side navigation pane. Unzip the file in a chosen location on your computer, e.g., the Downloads folder, and install the package from source.  

```{r install from source}
# install `charite` from tarball
# set your path
install.packages("~/Downloads/charite_0.3.0.tar.gz", repos = NULL, type = "source")
```

## Working on a computer with proxies  

`charite` requires packages `ggplot2`, `svglite`, `grid`, and `grDevices` to be installed first. To install them from CRAN, simply use `install.packages("packageName")`. To install R packages on a machine behind a proxy, you need to configure the proxy settings so R can connect to CRAN or GitHub through it.  

```{r install dependecies}
# create or write to existing .Rprofile
path2rprofile <- file.path(Sys.getenv("HOME"), ".Rprofile")
file.edit(path2rprofile)

# set proxies
# if your proxy requires authentication, use username:password@proxy.charite.de:8080
cat('Sys.setenv(http_proxy = "http://proxy.charite.de:8080")\n',
    'Sys.setenv(https_proxy = "http://proxy.charite.de:8080")\n',
    file = rprofile_path,
    append = TRUE)

# reload the .Rprofile settings
source(rprofile_path)
```

Once set, in RStudio go to Tools > Global Options > Packages and untick "Use Internet Explorer library/proxy for HTTP" and restart your R session. Note, this option may no longer be available in the latest versions of R/RStudio. Instead, try to download and install packages using `wininet` (Windows only) or `libcurl`.  

```{r install dependecies cont.}
# use "wininet" (a Windows-specific internet method) when installing a package
install.packages("packageName", method = "wininet")

# make this a permanent option by saving to your .Rprofile
cat('options(download.file.method = "wininet")\n', 
    file = file.path(Sys.getenv("HOME"), ".Rprofile"),
    append = TRUE)
```